\section{Conclusion and future work}

We have implemented a Coq plugin, which uses fast and lightweight \texttt{egg} Rust library, to automatically prove relational equivalences in Coq. We have developed a framework for communication between Coq and Rust and made a rule-parameterized algorithm on top of the \texttt{egg} library for producing a series of rewrites that is used in Coq interactive proofs. Moreover, we have experimented on various proving strategies and rewriting systems to implement an efficient and useful tool. 

We have defined multiple tools inside our plugin, that are used to interact with the \texttt{egg} library: 
\begin{itemize}
    \item \texttt{Cegg solve} --- a tactic that simplifies the lhs of the equation. It takes the conclusion of the goal, retrieves the lhs of the equivalence, parses it into an s-expression and passes it to \texttt{egg}. \texttt{Egg} builds an e-graph $E$ for it and extracts the ``best'' term $t$ from $E$. A sequence of rewrites to achieve $t$ is passed back to OCaml and is applied to the lhs of the equivalence inside Coq proof mode. 
    \item \texttt{Cegg solve eq} --- a tactic that tries to prove the equivalence between the lhs and rhs of the equation, using \texttt{egg}. 
    \item \texttt{Cegg config} --- a command, that allows to configure the ruleset for \texttt{egg}. It takes a user-defined list of rewrite rules and caches it for the later use in \texttt{Cegg solve} and \texttt{Cegg solve eq}.
\end{itemize}

As part of our future work, we plan to generalize our solution and extend its applicability to solve similar problems encountered by the \texttt{autorewrite} tactic in Coq. It takes a library of theorems as input and automatically tries to use it in order to prove the goal. Equality saturation has potential benefits and can outperform the existing approaches.  

The source code and documentation for the thesis is available at:
\begin{center}
    \href{https://github.com/K-dizzled/relations-via-egg}{\texttt{https://github.com/K-dizzled/relations-via-egg}}
\end{center}